---
title: "Reproducible Research _ Project1"
author: "López"
date: "25/8/2021"
output: html_document
---

Loading and preprocessing the data

```{r}
getwd()
direccion="C:/Users/user/Downloads/repdata_data_activity.zip"
unzip(direccion,exdir = "./data_activity")
```

Once we got the data already unziped, then we can proceed to read the data in a table o data frame format:

```{r}
activity_data=read.csv("./data_activity/activity.csv")
head(activity_data,5)
```

# Question 1, 2 ,3

Now we have the 3 variables dataset, thus we can calculate mean and median of steps per day:

- Including N.A
```{r}
total_steps = with(activity_data, tapply(steps, date, sum, na.rm = TRUE))
mean(total_steps)
median(total_steps)
```

-Not including N.A

```{r}
total_steps_na = with(activity_data, tapply(steps, date, sum, na.rm = FALSE))
mean(total_steps_na,na.rm = TRUE)
median(total_steps_na,na.rm = TRUE)
```

Or we get the same results working with tables:

-Including N.A
```{r}
activity_data_table=data.table::fread(input="./data_activity/activity.csv")
head(activity_data_table,5)
total_steps_tab = activity_data_table[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols = c("steps"), by = .(date)] 
total_steps_tab[, .(Mean = mean(steps, na.rm = TRUE), Median = median(steps, na.rm = TRUE))]
```

-Not including N.A

```{r}
total_steps_tab_na = activity_data_table[, c(lapply(.SD, sum, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)]

total_steps_tab_na[, .(Mean = mean(steps, na.rm = TRUE), Median = median(steps, na.rm = TRUE))]
```

And now constructing a histogram with the total number of steps taken per day and its frecuencies:

```{r warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(total_steps_tab, aes(x=as.Date(date),y=steps)) +
    geom_bar(stat = "identity") +
    labs(title = "Daily Steps", x = "Date", y = "Steps Frecuency")+
    theme_bw()

```

```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(hrbrthemes)
ggplot(total_steps_tab, aes(steps)) +
    geom_histogram(color="#e9ecef", alpha=0.6, position = 'identity')+
    labs(title = "Daily Steps", x = "Steps", y = "Steps Frecuency")+
        theme_bw()

```


From now on we will work just with tables for question 5 and 6.

# Question 4

Time series plot of the average number of steps taken:
```{r warning=FALSE, message=FALSE}
mean_steps_tab = activity_data_table[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)] 
head(mean_steps_tab,5)
library(ggplot2)
ggplot(mean_steps_tab, aes(x=as.Date(date),y=steps)) +
    geom_line(stat = "identity",color="black") +
    labs(title = "Avg. Daily Steps", x = "Date", y = "Avg. Steps Frecuency")+
    theme_bw()

```

As can be seen there are some missing data about specific dates, but this can be solve if we work with intervals, we solve this out in the next question

# Question 5

The 5-minute interval that, on average, contains the maximum number of steps:

```{r warning=FALSE, message=FALSE}
interval_5_min = activity_data_table[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 
head(interval_5_min,5)
ggplot(interval_5_min, aes(x = interval , y = steps)) + geom_line(color="black") + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")

```

And for the 5 minutes interval that contains the maximun number of steps on average:
```{r warning=FALSE, message=FALSE}
interval_5_min[steps == max(steps), .(max_interval = interval)]

```

# Question 6

Code to describe and show a strategy for imputing missing data:

```{r warning=FALSE, message=FALSE}
n_NA= nrow(activity_data_table[is.na(steps),])
# Or alternatively
n_NA_A=activity_data_table[is.na(steps),.N]
# Now that we determined the number os NA in the datasets, we need to propose an imputting method, for instance we'll use the mean
activity_data_table[is.na(steps), "steps"] = activity_data_table[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps")]

```

Once we got the imputed data we can proceed with the next questions.
# Question 7

Histogram of the total number of steps taken each day after missing values are imputed:

Before aswering this questions, we don´t want to overwrite the dataset, so let´s create a new one
```{r warning=FALSE, message=FALSE}
data.table::fwrite(x = activity_data_table, file = "./data_activity/imputed_activity.csv", quote = FALSE)
```

```{r warning=FALSE, message=FALSE}
mean_steps_tab = activity_data_table[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)] 
head(mean_steps_tab,5)
library(ggplot2)
ggplot(mean_steps_tab, aes(x=as.Date(date),y=steps)) +
    geom_line(stat = "identity",color="black") +
    labs(title = "Avg. Daily Steps", x = "Date", y = "Avg. Steps Frecuency")+
    theme_bw()

```

```{r warning=FALSE, message=FALSE}
total_steps_tab_na = activity_data_table[, c(lapply(.SD, sum, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)]
total_steps_tab_na[, .(Mean = mean(steps, na.rm = TRUE), Median = median(steps, na.rm = TRUE))]
ggplot(total_steps_tab, aes(steps)) +
    geom_histogram(color="#e9ecef", alpha=0.6, position = 'identity')+
    labs(title = "Daily Steps", x = "Steps", y = "Steps Frecuency")+
        theme_bw()

```

```{r warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(total_steps_tab_na, aes(x=as.Date(date),y=steps)) +
    geom_bar(stat = "identity") +
    labs(title = "Daily Steps", x = "Date", y = "Steps Frecuency")+
    theme_bw()

```


# Question 8

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:


```{r warning=FALSE, message=FALSE}
activity_data_table[, day := as.Date(date, format = "%Y-%m-%d")]
activity_data_table[, day:= weekdays(x = day)]
unique(activity_data_table$day)
activity_data_table[grepl(pattern = "lunes|martes|miércoles|jueves|viernes", x = day), "day_date"] = "weekday"
activity_data_table[grepl(pattern = "sábado|domingo", x = day), "day_date"] = "weekend"


#Now it's time to calcule the mean
interval_5_min_2 = activity_data_table[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, day_date)]
head(interval_5_min_2)

#Now it's time to make a plot 
library(ggplot2)
ggplot(interval_5_min_2, aes(x=interval,y=steps,colour=day_date)) +
    geom_line() +
    labs(title = "Avg. Daily Steps per 5_min Interval on Weekdays and Weekends", x= "Interval",y= "Avg. Steps Frecuency")+
    facet_wrap(~day_date , ncol = 1, nrow=2)+
    theme_bw()

```
